The effect of detector nonlinearity on \wfa\ PSF profiles for weak gravitational lensing measurements

### 2016-04-25 ms.tex
Edited section about dithering.  Tweaked figure captions

To illustrate the effect of NL on photometry, consider nearly saturated stars in the HLS.  Using PSF profiles with a simulated AB magnitude up to 18.3 for an exposure time of 168.1 seconds (expected for the HLS), the total flux in each band is shown in Table 1, along with the peak pixel value. When the profile is drawn on a pixelated postage stamp with a scale of $p=0.11$ arcseconds and placed on the center of the pixel\footnote{Note that the peak value is a function of the PSF profile centroid location within the pixel, as well as pixel resolution. When drawing the PSF profiles in our simulations, we randomize the PSF centroid within the native scale pixel, as described in Section 3.}, this particular magnitude produces a peak charge of $\sim 1\times 10^5$ e$^{-}$ (Y106 band, see Table 1), which represents about 90\% of the typical pixel full well value of $1.1 \times 10^5$ e$^{-}$.\footnote{Dave Content, private communication.}
At this level of charge, the signal attenuation due to NL for the nominal $\beta_0$ value is of about 5\% (Eqn. \ref{NL}).

@@

To overcome this problem in real data, multiple dithered exposures are taken, and then processed by an image combination algorithm (\citealt{lauer99b}, \citealt{fruchter02}, \citealt{bertin06}, \citealt{rowe11}) in order to produce an oversampled image that satisfies the Nyquist-Shannon criterium. To study NL with simulations, however, we do not need to make use of such external software\textemdash which can be computationally expensive\textemdash to produce an image at higher resolution.  Instead, we render the convolved PSF profile\textemdash including the pixel response at its native scale\textemdash as a \gs\ object at a high resolution through the method {\tt{galsim.GSObject.drawImage}},\footnote{The keyword {\tt{method}} should be set to ``{\tt{no\char`_pixel}}" when calling {\tt{galsim.GSObject.drawImage}}, since by default it convolves its argument profile with a boxcar of length {\tt{pixel\char`_scale}}. This extra convolution must be turned off because the effective PSF profile we create already includes this pixel response.} setting the parameter {\tt{scale}} to $p/N$, where $N$ is a positive integer and $p$ is the native pixel scale.

Defining $Q\equiv1/p\times u_{\text{max}}$ as the sampling factor ($Q<2$, $Q=2$, and $Q\geq2$ represent under-, critically, and over-sampled images, respectively), we see that in order to produce an over-sampled image ($Q\geq2$), the Nyquist-Shannon criterium implies that $N$ must be given by (see \emph{e.g.}, \citealt{marks09}, \citealt{shapiro13})

\begin{align}
N=\frac{2p}{\lambda_{min} F}
\label{nimages}
\end{align}
where $p$ is the pixel size, $\lambda_{min}$ is the shortest wavelength in a given filter, and $F$ is the the f-number of the telescope. In the case of the \wfa\ telescope, $p=10\ \mu$m and $F=7.8$, resulting in $N\geq3$ ($N\geq2$) for the J129 (H158) band. Thus, we have chosen to set $N=3$ in our simulations
%, which will produce a number of images similar to what the WL survey in \wf\ will use per exposure (5\textendash7 across two epochs). 
to ensure oversampling.
At this resolution, it will also be possible to study the impact of centroid shifts within the native pixels (since NL depends on the absolute pixel values it will also depend on where the star is centered within a particular pixel), and therefore for our simulations we randomly vary the PSF centroid within a native pixel. \textbf{[WHAT DOES THIS MEAN?]}

@@

The PSF profiles are drawn into squared postage stamps of size $1.5k$ pixels ($k=64$), large enough to ensure that most ($\sim 96\%$) of the total PSF flux is contained in the stamp. %In addition, this avoid biases in PSF determination arising from object selection and deblending. 
Each profile object was assigned a total flux of

@@

Notice that in our simulations we have neglected the main sources of noise that would affect the HLS\textemdash zodiacal background, thermal emission, and read noise\textemdash which would make all images slightly more nonlinear. Calculations performed with the \wfa\ Exposure Time Calculator v.14\footnote{\url{https://wfirst.ipac.caltech.edu/sims/tools/wfDepc/wfDepc.html}} \citep{hirata12} in weak lensing mode (ETC-WL) by \citealt{spergel15} in the F184 band (a conservative case) show that the combined contribution due to these backgrounds sources would be approximately 130 $e^-$ per pixel for a $174$ seconds exposure (comparable to the $168.1$ exposure time assumed in this work), negligible compared with the values used in our simulations (Table 1).

@@

\begin{figure}[!h]
\centering
\resizebox{\hsize}{!}{\includegraphics[page=1]{plots/f2.eps}}
\caption{Fractional error in PSF size (upper panel) and absolute error in PSF ellipticity components (lower panels) as a function of the mean nonlinearity parameter $\beta$ for the \wfa\ PSF, in four the four HSL filters (J129, Y106, H158, and F184), and an AB magnitude of 18.3 at 168.1 seconds of exposure time (similar plots were found for other magnitude values). Each point is the mean over 100 realizations of uniformly distributed random centroid shifts within the high resolution images ($N=3$), and the associated error bars (too small to be seen) represent the standard deviation of the realizations.}
\label{f2}
\end{figure}

@@

The trends in Fig. \ref{f3} are clearly well-described by a power law with a common slope for the various filters.  We fit a function only to the PSF size trend since it will be a more robust characteristic of the  survey, whereas ellipticity can have relatively wide spatial variations.  For a power law of the form
\begin{align}
\frac{\Delta R/ R}{\beta} = A_{\text{F}} 10^{B_{\text{F}} (m - m_{0})} \ (m_0 =20), 
\label{power_law}
\end{align}
the parameters $A_{\text{F}}$ and $B_{\text{F}}$ for each filter F are listed in Table 2.

@@

\begin{figure}[!h]
\centering
\resizebox{\hsize}{!}{\includegraphics[page=1]{plots/f4.eps}}
\caption{Impact of spatial variability in the $\beta_0$ parameter, quantified as the standard deviation over $M=100$ realizations (for a fixed centroid) in $\Delta e_1$,  $\Delta e_2$, and $\Delta R/R$  as a function of the standard deviation $ \sigma_{\beta_0}$. }
\label{f4}
\end{figure}
We found that the results in Fig. \ref{f4} are insensitive to the nominal $\beta$ and therefore, in general, $\sigma_{\beta}$  can be treated as an error on $\beta$ estimation. The linear relationships in Fig. \ref{f4} also allows us to plot the slope of each curve for different magnitudes, as was done in Fig. \ref{f3}, and therefore analogous equations to Eqn. \ref{calib} can be used to derive requirements on the maximum tolerable dispersion in size and ellipticity errors given a particular set of parameters and requirements. The results are shown in Fig. \ref{f5}. We also fitted a model of the form 
\begin{align}
\frac{\sigma_{\Delta R/ R}} {\sigma_{\beta}} = A_{\text{F}}10^{B_{\text{F}} (m - m_{0})} \ (m_0 =20)
\end{align}
for the PSF size trends in Fig. \ref{f5}, as was done above for Fig. \ref{f3} (\emph{c.f.}, Eqn. \ref{power_law}). The fit parameters are presented in Table 2 as well.

%Table 2
\begin{table}
\centering
\begin{tabular}{|c|c|c|c|c|}
\hline
\multirow{2}{*}{Band} & \multicolumn{2}{c|}{size bias (Fig. \ref{f3})} & \multicolumn{2}{c|}{size scatter (Fig. \ref{f5})} \\
\cline{2-5}
& $A_{\text{F}}$ & $B_{\text{F}}$ & $\ A_{\text{F}}$ & $B_{\text{F}}$  \\
\hline
Y106 & 8322 & -0.4086 & 5057 &  -0.4045  \\
\hline
J129 & 6958 & -0.4089  & 4215 &  -0.4050 \\
\hline
H158 & 4668 & -0.4069 & 2711 &  -0.4038 \\
\hline
F184 & 2209 & -0.4035 & 1194 &  -0.4018 \\
\hline
\end{tabular}
\caption{Parameters resulting from fitting the curves in the upper panel of Figs. \ref{f3} and \ref{f5} to power-law functions of the form $\Delta R/ R/ \beta = A_{\text{F}}\ 10^{B_{\text{F}} (m - 20)}$ and $\sigma_{\Delta R/ R}/ \sigma_{\beta} = A_{\text{F}}\ 10^{B_{\text{F}} (m - 20)}$, respectively.}
\label{table2}
\end{table}

\begin{figure}[!h]
\centering
\resizebox{\hsize}{!}{\includegraphics[page=1]{plots/f5.eps}}
\caption{Standard deviation of the fractional error in PSF size and of the absolute error in PSF ellipticity components normalized by the NL model parameter $\beta$, as a function of magnitude for the four HSL filters (J129, Y106, H158, and F184).}
\label{f5}
\end{figure}

>>>>

To illustrate the effect of NL on photometry, consider nearly saturated stars in the HLS.  Using PSF profiles with a simulated AB magnitude up to 18.3 for an exposure time of 168.1 seconds (expected for the HLS), the total flux in each band is shown in Table 1, along with the peak pixel value. When the profile is drawn on a pixelated postage stamp with a scale of $p=0.11$ arcseconds and placed on the center of the pixel\footnote{Note that the peak value is a function of the PSF profile centroid location within the pixel, as well as pixel resolution. When drawing the PSF profiles in our simulations, we randomize the PSF centroid within the native scale pixel, as described in Section 3.}, this particular magnitude produces a peak charge of $\sim 1\times 10^5$ e$^{-}$ (Y106 band, see Table 1), which represents about 90\% of the typical pixel full well value of $1.1 \times 10^5$ e$^{-}$.\footnote{Dave Content, private communication.}
At this level of charge, the signal attenuation due to NL for the nominal $\beta_0$ value is about 5\% (Eqn. \ref{NL}).

@@

To overcome this problem in real data, multiple dithered exposures are taken, and then processed by an image combination algorithm (\citealt{lauer99b}, \citealt{fruchter02}, \citealt{bertin06}, \citealt{rowe11}) in order to produce an oversampled image that satisfies the Nyquist-Shannon criterium. To study NL with repeated simulations, however, such external software is computationally expensive.  Instead, we approximate oversampled data by rendering the convolved PSF profile\textemdash including the pixel response at its native scale\textemdash as a \gs\ object at a high resolution through the method {\tt{galsim.GSObject.drawImage}},\footnote{The keyword {\tt{method}} should be set to ``{\tt{no\char`_pixel}}" when calling {\tt{galsim.GSObject.drawImage}}, since by default it convolves its argument profile with a boxcar of length {\tt{pixel\char`_scale}}. This extra convolution must be turned off because the effective PSF profile we create already includes this pixel response.} setting the parameter {\tt{scale}} to $p/N$, where $N$ is a positive integer and $p$ is the native pixel scale. 
 Defining $Q\equiv1/p\times u_{\text{max}}$ as the sampling factor ($Q<2$, $Q=2$, and $Q\geq2$ represent under-, critically, and over-sampled images, respectively), we see that in order to produce an over-sampled image ($Q\geq2$), the Nyquist-Shannon criterium implies that $N$ must be given by (see \emph{e.g.}, \citealt{marks09}, \citealt{shapiro13})

\begin{align}
N=\frac{2p}{\lambda_{min} F}
\label{nimages}
\end{align}
where $p$ is the pixel size, $\lambda_{min}$ is the shortest wavelength in a given filter, and $F$ is the the f-number of the telescope. In the case of the \wfa\ telescope, $p=10\ \mu$m and $F=7.8$, resulting in $N\geq3$ ($N\geq2$) for the J129 (H158) band. Thus, we have chosen to set $N=3$ in our simulations
%, which will produce a number of images similar to what the WL survey in \wf\ will use per exposure (5\textendash7 across two epochs). 
to ensure oversampling.

Note that the NL effect on the reconstructed image from dithering real exposures will depend on the centroid of the source in the dithered inputs: an exposure where the PSF peaks at a pixel center will have higher NL than an exposure where the peak is spread more evenly over 2-4 pixels.  To mimic this behavior in our approximation, we randomly offset our PSF centroids 100 times uniformly over a box the size of a native detector pixel, and we average all shape measurements over these realizations.  With $N$=3, the variation of the NL correction in the pixels of the oversampled image should be similar to the variations in the 5-8 exposures combined using the WFIRST dither strategy.  In practice, we find that the scatters of the shape measurements over these realizations are negligible relative to the means.

%At this resolution, it will also be possible to study the impact of centroid shifts within the native pixels (since NL depends on the absolute pixel values it will also depend on where the star is centered within a particular pixel), and therefore for our simulations we randomly vary the PSF centroid within a native pixel. \textbf{[WHAT DOES THIS MEAN?]}

@@

The PSF profiles are drawn into squared postage stamps of size $1.5k$ pixels ($k=64$), large enough to ensure that most \textbf{[HOW MUCH]} of the total PSF flux is contained in the stamp. %In addition, this avoid biases in PSF determination arising from object selection and deblending. 
Each profile object was assigned a total flux of

@@

We have neglected the main sources of noise that would affect the HLS\textemdash zodiacal background, thermal emission, and read noise\textemdash which would make all images slightly more nonlinear. Calculations performed with the \wfa\ Exposure Time Calculator v.14\footnote{\url{https://wfirst.ipac.caltech.edu/sims/tools/wfDepc/wfDepc.html}} \citep{hirata12} in weak lensing mode (ETC-WL) by \citealt{spergel15} in the F184 band (a conservative case) show that the combined contribution due to these backgrounds sources would be approximately 130 $e^-$ per pixel for a $174$ seconds exposure (comparable to the $168.1$ exposure time assumed in this work), which creates negligible corrections in our simulations for a NL parameter $\beta$ of order $10^{-6}$.

@@

\begin{figure}[!h]
\centering
\resizebox{\hsize}{!}{\includegraphics[page=1]{plots/f2.eps}}
\caption{Fractional error in PSF size (upper panel) and absolute error in PSF ellipticity components (lower panels) as a function of the mean nonlinearity parameter $\beta$ for the \wfa\ PSF, in four the four HSL filters (J129, Y106, H158, and F184), and an AB magnitude of 18.3 at 168.1 seconds of exposure time. 
Each point is the mean over 100 realizations of uniformly distributed random centroid shifts within the high resolution images ($N=3$).
The standard deviations of the realizations are negligible.}
\label{f2}
\end{figure}

@@

The trends in Fig. \ref{f3} are clearly well-described by a power law with a common slope for the various filters.  We fit a function only to the PSF size trend since it will be a more robust characteristic of the  survey, whereas ellipticity can have relatively wide spatial variations.  For a power law of the form $\Delta R/ R/ \beta = A_{\text{F}} 10^{B_{\text{F}} (m - m_{0})}$ (for $m_0 =20$), the parameters $A_{\text{F}}$ and $B_{\text{F}}$ for each filter F are listed in Table 2.

@@

\begin{figure}[!h]
\centering
\resizebox{\hsize}{!}{\includegraphics[page=1]{plots/f4.eps}}
\caption{Impact of spatial variability in the $\beta_0$ parameter, quantified as the standard deviation over $M=100$ realizations (for a fixed centroid) in $\Delta e_1$,  $\Delta e_2$, and $\Delta R/R$  as a function of the standard deviation $ \sigma_{\beta_0}$. 
The source magnitude is fixed at 18.3.}
\label{f4}
\end{figure}
We found that the results in Fig. \ref{f4} are insensitive to the nominal $\beta$ and therefore, in general, $\sigma_{\beta}$  can be treated as an error on $\beta$ estimation. The linear relationships in Fig. \ref{f4} also allows us to plot the slope of each curve for different magnitudes, as was done in Fig. \ref{f3}, and therefore analogous equations to Eqn. \ref{calib} can be used to derive requirements on the maximum tolerable dispersion in size and ellipticity errors given a particular set of parameters and requirements. The results are shown in Fig. \ref{f5}. We also fitted a model of the form \textbf{[MAKE NUMBERED EQ]} $\sigma_{\Delta R/ R}/ \sigma_{\beta} = A_{\text{F}}10^{B_{\text{F}} (m - m_{0})}$ (for $m_0 =20$) for the PSF size trends in Fig. \ref{f5}, as was done above for Fig. \ref{f3}. The fit parameters are presented in Table 2 as well.

%Table 2
\begin{table}
\centering
\begin{tabular}{|c|c|c|c|c|}
\hline
\multirow{2}{*}{Band} & \multicolumn{2}{c|}{size bias (Fig. \ref{f3})} & \multicolumn{2}{c|}{size scatter (Fig. \ref{f5})} \\
\cline{2-5}
& $A_{\text{F}} (\times 10^3)$ & $B_{\text{F}}$ & $\ A_{\text{F}} (\times 10^3)$ & $B_{\text{F}}$  \\
\hline
Y106 & 8.322 & -0.4086 & 5.057 &  -0.4045  \\
\hline
J129 & 6.958 & -0.4089  & 4.215 &  -0.4050 \\
\hline
H158 & 4.668 & -0.4069 & 2.711 &  -0.4038 \\
\hline
F184 & 2.209 & -0.4035 & 1.194 &  -0.4018 \\
\hline
\end{tabular}
\caption{Parameters resulting from fitting the curves in the upper panel of Figs. \ref{f3} and \ref{f5} to power-law functions of the form $\Delta R/ R/ \beta = A_{\text{F}}\ 10^{B_{\text{F}} (m - 20)}$ and $\sigma_{\Delta R/ R}/ \sigma_{\beta} = A_{\text{F}}\ 10^{B_{\text{F}} (m - 20)}$, respectively. \textbf{[THE x$10^3$ is confusing -- just use whole numbers?]}}
\label{table2}
\end{table}

\begin{figure}[!h]
\centering
\resizebox{\hsize}{!}{\includegraphics[page=1]{plots/f5.eps}}
\caption{Standard deviation of the fractional error in PSF size and of the absolute error in PSF ellipticity components normalized by the error in the NL parameter $\sigma_\beta$, as a function of magnitude for the four HSL filters (J129, Y106, H158, and F184).}
\label{f5}
\end{figure}
